True versus apparent shapes of bow shocks

### 2018-03-05 papers/Papers1234/sec-conclusions.tex
Finish discussion of systematic uncertainties

@@

We determine the planitude and alatude by fitting a circle to the
traced points within \(\pm \Delta\theta = \ang{75}\) bow axis, using the
iterative algorithm described in Appendix~\ref{app:rcurv-empirical}.
The fitted circle, when combined with the position of the central
source, yields the orientation of the bow axis, together with the apex
distance, \(R_0\), radius of curvature, \(R_\C\), and two
perpendicular radii (one for each wing), \(R_{90+}\) and \(R_{90-}\).
These are all indicated on the panels of Figure~\ref{fig:000-400-fit}
by light-colored lines.\footnote{%
  For conciseness, we drop the prime symbol from the radii, both in
  this section and in Appendix~\ref{app:rcurv-empirical}, even though
  they are all projected quantities.} %
The projected planitude and alatude then follow as \(\Pi' = R_\C/R_0\),
\(\Lambda_+ = R_{90+}/R_0\), \(\Lambda_- = R_{90-}/R_0\), which are shown in
Figure~\ref{fig:000-400-planitude-alatude}.

The planitude is found to have a moderate dependence on the choice of
\(\Delta\theta\), as shown in Figure~\ref{fig:000-400-Delta-theta}, where it can
be seen that, although the values of \(\Pi'\) are relatively stable for
\(\Delta\theta \ge \ang{60}\) (light gray shaded region), they can show much
larger variations for \(\Delta\theta < \ang{60}\).  The fact that the radius of
curvature is defined at a point (the projected apex) might seem to
argue for making \(\Delta\theta\) as small as possible, but that would lead to
circle fits that were extremely sensitive to the exact positions of
the few points included in the fit.  A reliable fit requires 4 or more
points, ideally spanning a total separation that is a substantial
fraction of \(R_\C\), which would argue for \(\Delta\theta\) larger than about
\(\Pi'/2\) radians, or \ang{60}--\ang{90}.  On the other hand, if
\(\Delta\theta \ge \ang{90}\) then the planitude and alatude would no longer be
independent since the bow would be forced to lie on the ``sphere''
line, \(\Lambda' = (2 \Pi' - 1)^{1/2}\) (see \S~\ref{sec:conic}).  Balancing
these two concerns suggests an optimal \(\Delta\theta = \ang{75}\), which is
shown in Figure~\ref{fig:000-400-fit}, whereas in
Figure~\ref{fig:000-400-planitude-alatude} we show results for both
\ang{75} (thick lines) and \ang{60} (thin lines).

Unlike all the models considered in \S~\ref{sec:crw-scenario} and
\S~\ref{sec:more-realistic-bow}, the observed bows are not necessarily
symmetrical and so the alatude for the two wings, \(\Lambda_-'\) and
\(\Lambda_+'\), may be different.  We calculate an average alatude,
\(\langle\Lambda'\rangle\), and an alatude asymmetry, \(\Delta\Lambda'\):
\begin{equation}
  \label{eq:alatude-average-and-asym}
  \langle\Lambda'\rangle = \tfrac12 \left( \Lambda_+' + \Lambda_-' \right)
  \quad\quad \Delta\Lambda' = \tfrac12 \left( \Lambda_+' - \Lambda_-' \right)
\end{equation}
The results for these and the planitude are shown in Table~\ref{tab:m42-fit}

\(\Pi' = 2.9 \pm 0.2 \pm 0.3\)

>>>>

\subsection{Empirical determination of bow shock shape}
\label{sec:empir-determ-bow}

@@

We determine the planitude and alatude by fitting a circle to the
traced points within \(\pm \Delta\theta = \ang{75}\) of the bow axis, using the
iterative algorithm described in Appendix~\ref{app:rcurv-empirical}.
The fitted circle, when combined with the position of the central
source, yields the orientation of the bow axis, together with the apex
distance, \(R_0\), radius of curvature, \(R_\C\), and two
perpendicular radii (one for each wing), \(R_{90+}\) and \(R_{90-}\).
These are all indicated on the panels of Figure~\ref{fig:000-400-fit}
by light-colored lines.\footnote{%
  For conciseness, we drop the prime symbol from the radii, both in
  this section and in Appendix~\ref{app:rcurv-empirical}, even though
  they are all projected quantities.} %
The projected planitude and alatude then follow as
\(\Pi' = R_\C/R_0\), \(\Lambda_+ = R_{90+}/R_0\),
\(\Lambda_- = R_{90-}/R_0\), which are shown in
Figure~\ref{fig:000-400-planitude-alatude}.

\subsection{Analysis of sources of systematic uncertainty}
\label{sec:analys-sourc-syst}

The planitude is found to have a moderate dependence on the choice of
\(\Delta\theta\), as shown in Figure~\ref{fig:000-400-Delta-theta}, where it can
be seen that, although the values of \(\Pi'\) are relatively stable for
\(\Delta\theta \ge \ang{60}\) (light gray shaded region), they can show much
larger variations for \(\Delta\theta < \ang{60}\).  The fact that the radius of
curvature is defined at a point (the projected apex) might seem to
argue for making \(\Delta\theta\) as small as possible, but that would lead to
circle fits that were extremely sensitive to the exact positions of
the few points included in the fit.  A reliable fit requires 4 or more
points, ideally spanning a total separation that is a substantial
fraction of \(R_\C\), which would argue for \(\Delta\theta\) larger than about
\(\Pi'/2\) radians, or \ang{60} to \ang{90}.  On the other hand, if
\(\Delta\theta \ge \ang{90}\) then the planitude and alatude would no longer be
independent since the bow would be forced to lie on the ``sphere''
line, \(\Lambda' = (2 \Pi' - 1)^{1/2}\) (see \S~\ref{sec:conic}).  Balancing
these two concerns suggests an optimal \(\Delta\theta = \ang{75}\), which is
shown in Figure~\ref{fig:000-400-fit}, whereas in
Figure~\ref{fig:000-400-planitude-alatude} we show results for both
\ang{75} (thick lines) and \ang{60} (thin lines).

Unlike all the models considered in \S~\ref{sec:crw-scenario} and
\S~\ref{sec:more-realistic-bow}, the observed bows are not necessarily
symmetrical and so the alatude for the two wings, \(\Lambda_-'\) and
\(\Lambda_+'\), may be different.  We therefore calculate an average
alatude, \(\langle\Lambda'\rangle\), and an alatude asymmetry, \(\Delta\Lambda'\):
\begin{equation}
  \label{eq:alatude-average-and-asym}
  \langle\Lambda'\rangle = \tfrac12 \left( \Lambda_+' + \Lambda_-' \right)
  \quad\quad \Delta\Lambda' = \tfrac12 \left( \Lambda_+' - \Lambda_-' \right)
\end{equation}
The results for these two quantities, together with the planitude, are
shown in Table~\ref{tab:m42-fit}.  For each object and for each tracing
(inner, ridge, outer, see Fig.~\ref{fig:000-400-fit}), the
\(\text{mean} \pm \text{standard\ deviation}\) is listed for circle
fits using \(\Delta\theta = \ang{60}\) to \ang{80} (see
Fig.~\ref{fig:000-400-Delta-theta}).  Additionally, the row ``All''
for each object gives the global mean and standard deviation over all
three tracings.

It can be seen from Table~\ref{tab:m42-fit} that the uncertainty in
the fitted parameters is dominated by the variations between the
different traced shapes.  For example, the one-sigma relative
variation of planitude \(\Pi'\) is \(< 10\%\) within the individual
tracings, but \(\approx 20\%\) between tracings.  For the mean alatude
\(\langle\Lambda'\rangle\), the variation within individual tracings is extremely
small\footnote{%
  This is because the only way that variation in the circle fit
  parameters affect the alatudes is through the axis orientation
  vector, and changing the orientation induces roughly opposite
  effects on \(\Lambda_+'\) and \(\Lambda_-'\), which cancel out in \(\langle\Lambda'\rangle\).
} %
at \(\approx 1\%\), but is \(\approx 10\%\) between tracings.  The alatude
asymmetry, \(\Delta\Lambda'\), is best interpreted as a difference between the
symmetry axis of the apex region and the symmetry axis of the wings.
In relative terms, this is \(0\)--\(20\%\), but with large systematic
variations between tracings (for instance, in both objects it is very
small for the outer arcs).

It could be argued that much of the between-tracing variations in
\(\Pi'\) and \(\langle\Lambda'\rangle\) are due to real differences in the shape of the
inner and outer boundaries of the emission arc.  Although this may be
true, in the absence of a robust theory as to exactly what feature of
the observed images constitutes \emph{the} bow shock, such variations
nevertheless serve to limit the precision with which the bow shock
shape can be measured.  We therefore conclude that conservative
estimates of \(20\%\) precision for \(\Pi'\) and \(10\%\) precision for
\(\Lambda'\) are appropriate when analyzing observations of a similar
quality to those presented in Figure~\ref{fig:000-400-fit}.

\subsection{Derived shape of M42 arcs}
\label{sec:derived-shape-m42}
